Refunctionalization of Abstract Abstract Machines (Functional Pearl)

### 2018-06-24 paper/paper.tex
a small revision
fix code

Traditionally, continuation-passing style (CPS) is a popular intermediate representation
for analyzing functional programs because it exposes control transfer explicitly and 
simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
Here, we choose to use $\lambda$-calculus in administrative normal form (ANF)
\cite{flanagan1993essence}, which is a direct-style intermediate representation, as our 
target for clarity, but without losing simplicity and generality. The transformations 
we will show in the rest of this paper also work on abstract machines for plain 
direct-style $\lambda$-calculus languages. Although we only show the core calculus language,
it can be easily extended to support recursive bindings (such as @letrec@), conditionals, 
primitive types, and operations on primitive types. These cases would be trivial to implement,
so we elide them in this paper.

@@

\begin{lstlisting}
  case class Store[K,V](map: Map[K, Set[V]]) {
    def apply(addr: K): List[V] = map(addr).toList
    def update(addr: K, d: Set[V]): Store[K,V] = {
      val oldd = map.getOrElse(addr, Set())
      Store[K, V](map ++ Map(addr -> (d ++ oldd)))
    }
    def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
  }
  type BStore = Store[BAddr, Storable];  type KStore = Store[KAddr, Cont]
\end{lstlisting}

>>>>

Traditionally, continuation-passing style (CPS) is a popular intermediate representation
for analyzing functional programs because it exposes control transfer explicitly and 
simplifies analysis~\cite{Shivers:1991:SSC:115865.115884, Shivers:1988:CFA:53990.54007}.
Here, we choose to use $\lambda$-calculus in administrative normal form (ANF)
\cite{flanagan1993essence}, which is a direct-style intermediate representation, as our 
target for clarity, but without losing simplicity and generality. The transformations 
we will show in the rest of this paper also work on abstract machines for plain 
direct-style $\lambda$-calculus languages. Although we only show the core calculus language,
it can be easily extended to support recursive bindings (such as @letrec@), conditionals, 
primitive types, and operations on primitive types. These cases would be straightforward to
implement without introducing new transformations as far as the concerns of this paper,
so we elide them here.
% TODO: appendix, direct-style, numbers and if0

@@

\begin{lstlisting}
  case class Store[K,V](map: Map[K, Set[V]]) {
    def apply(addr: K): Set[V] = map(addr)
    def update(addr: K, d: Set[V]): Store[K,V] = {
      val oldd = map.getOrElse(addr, Set())
      Store[K, V](map ++ Map(addr -> (d ++ oldd)))
    }
    def update(addr: K, sd: V): Store[K,V] = update(addr, Set(sd))
  }
  type BStore = Store[BAddr, Storable];  type KStore = Store[KAddr, Cont]
\end{lstlisting}
